---
title: "Iris Analysis"
author: "Ale F. Benages"
date: "3/22/2021"
output: html_document
---


# Libraries
```{r}
#install.packages('tidyverse')
#install.packages(readr)
library("readr")
library(ggplot2)
```

# Load CSV
```{r}
IrisDataset <- read.csv('iris.csv')
```

# 1. EDA 
## 1.1 Knowing your data
```{r}
# To list the attributes of the data set.
attributes(IrisDataset)
# To print the min, max, mean, median, and quartiles of each attribute.
summary(IrisDataset)
# To display the structure of your data set.
str(IrisDataset) 
# To get th names of the attributes within your data set.
names(IrisDataset) 
```

As the analysis goal is about the relationship between the Petal Length and Petal Width, let's analyze those variables detailed
```{r}
# specific info about the Petal.Lenght 
summary(IrisDataset$Petal.Length)
# specific info about the Petal.Width 
summary(IrisDataset$Petal.Width)
```



## 1.2 Missing Values?
If there are, the summarye() will count how many NA’s you have.
```{r}
summary(IrisDataset)
```

Also you can show your NA’s through logical data. (TRUE if it’s missing, FALSE if it’s not.)

```{r}
is.na(IrisDataset) 
```

Replace the missing values with the mean, which is common technique, but **something to use with care with as it can skew the data**.
```{r}
#DatasetName$ColumnName[is.na(DatasetName$ColumnName)]<-mean(DatasetName$ColumnName,na.rm = TRUE)
```

## 1.3 Some Plots
### PETAL'S LENGTH

```{r}
hist(IrisDataset$Petal.Length, 
     main="Sepal Length distribution", xlab="Lenght [mm]", 
     col='green')
```


```{r}
#' boxplot
ggplot(data = IrisDataset, aes(x = "", y = Petal.Length)) + 
  geom_boxplot() + theme_classic()+ 
  xlab("")+ylab("Petal Lenght [mm]")
```

In this case, Boxplot it's not being very descriptive, let's try a Violinplot:

```{r}
#' violinplot
ggplot(data = IrisDataset, aes(x = "", y = Petal.Length)) + 
  geom_violin() + theme_classic()+ 
  xlab("")+ylab("Petal Lenght [mm]")+ 
  geom_boxplot(width=0.1)+ 
  stat_summary(fun=median, geom="point", size=2, color="red")
```

### PETAL'S WIDTH

```{r}
hist(IrisDataset$Petal.Width, 
     main="Sepal Length distribution", xlab="Lenght [mm]", 
     col='violet')
```


```{r}
#' boxplot
ggplot(data = IrisDataset, aes(x = "", y = Petal.Width)) + 
  geom_boxplot() + theme_classic()+ 
  xlab("")+ylab("Petal Lenght [mm]")
```


```{r}
#' violinplot
ggplot(data = IrisDataset, aes(x = "", y = Petal.Width)) + 
  geom_violin() + theme_classic()+ 
  xlab("")+ylab("Petal Lenght [mm]")+ 
  geom_boxplot(width=0.1)+ 
  stat_summary(fun=median, geom="point", size=2, color="red")
```


### Petal's Length vs Petal's Width
```{r}
ggplot(data = IrisDataset, aes(x = Petal.Length  , y = Petal.Width))  + 
  geom_point(shape = 1) + 
  xlab("Petal Lenght [mm]")+ylab("Petal Width [mm]") +
  theme_bw()
```

There is a positive correlation between those variables. 




# 2. Testing and Training Sets

First of all, define a seed to get *reproducible experiments*. 
```{r}
set.seed(122)
```

These two lines calculate the sizes of each set **but do not create** the sets:
```{r}
trainSize<-round(nrow(IrisDataset)*0.8) # 80/20%
testSize<-nrow(IrisDataset)-trainSize
```

Just to check the dimension of the empty vectors.
```{r}
print(trainSize)
print(testSize)
```


So, how do you create the **train/test sets**?. We also want these sets to be in a randomized order, which will help to create an optimal model. To perform this, you need to run these three lines of code. 
```{r}
training_indices<-sample(seq_len(nrow(IrisDataset)),size =trainSize)
trainSet<-IrisDataset[training_indices,]
testSet <-IrisDataset[-training_indices,]
```


# 3. Linear Regression 
The next step is to construct a model to answer the business question: petal's length using the petal’s width. 

* **Input / Independent Variable / Y **: petal’s width
* **Output / Dependent Variable / X **: petal's length

To do this, we’ll be using the prediction function – predict() to get a **first grade** equation that fit's the data. 

```{r}
# The basic line of code for the linear model function. 
LinearModel<- lm(Petal.Width ~ Petal.Length, trainSet )
```

thats all!. Now to see the metrics of the model, just call **summary()**
```{r}
summary(LinearModel)
```


## 3.1 Predictions

To make predictions, we use the **Test Set**.
```{r}
prediction <- predict(LinearModel, testSet)
prediction
```
