---
title: "CARS - Report"
author: "Ale F. Benages"
date: "3/22/2021"
output: html_document
---


```{r}
print('a')



```


# 1 Description of the dataset
 
* Consists on 3 columns (‘Name of cars’, ‘Speed” and “Distance”), and 50 rows of records.  
* No Missing values.   

## 1.1 Variable: ‘Name of cars’  
* Nominal Categorical Variable.   
* Holds the different brands of cars. 

## 1.2 Variable: ‘Speed’
* Numerical variable
* As I didn’t have any extra information about the dataset, I assumed that this variables represent the speed of cars, and it’s units are Mph ( a max velocity of 25kmh is quite low, but 25 mph is more acceptable, around 40kmh). 
* This graph summarizes the numerical values of this variable: 

```{r}
summary(speed)
```

Speed
Min.   : 4.0   
1st Qu.:12.0   
Median :15.0   
Mean   :15.4  
3rd Qu.:19.0   
Max.   :25.0   

The most frequent speed of those cars are between 12 to 19 mph. And the average speed is 15.4 mph. 


    • ‘Distance’, Numerical Variable. 
        ◦ I assumed that this variable represents the stopping distance, and that the measuring unit is feet. (also those values have more sense in feet than meters because a stopping distance of 120m is quite exaggerated, but in feet it’s a safer distance of almost 37 meters) 
        ◦ This graph summarizes the numerical values of this variable. 
Distance       
Min.   :  2.00  
1st Qu.: 26.00  
Median : 36.00  
Mean   : 42.98  
3rd Qu.: 56.00  
Max.   :120.00  

The typical stopping distance of those cars are between 26 to 56 feet. And the average is 43 feet.
There is one car that has an exaggerated stopping distance of 120 feet.  


Useful information comes from plotting both variables together. There is a clear positive correlation  between them. This correlation confirms the assumption about the units and description of the variables: as faster a car goes, more distance will be needed to stop it.




Prediction
The dataset was divided in train/test set with a rate of 70/30.  And then a linear regression model was applied to achieve the goal of the analysis: predict the car stopping distance, based on tits speed.
    • Independent Variable: Speed of the cars (speed)
    • Dependent Variable: Stopping distances (distance)
And this data fits into the linear regression as:   Distance = slope*speed + y_intercept 
With the results, the equation of the model changes to   Distance = 3.582*speed – 13.082


Conclusion
A detailed output could be found in the output of the script. To summarize, I will focus on the result of the R2, as this statistic provides a measure of how well the model fits the data. Those values lie between 0 and 1 (where 0 represents a regression that does not explain the relation of the variables at all, and 1 corresponds to a full and perfect correlation). In our case we got a Multiple R-squared of 0.603. Our rough 60% of the changes in the response variable (Stopping Distance) could be explained by the predictor variable (Speed). And it also makes sense, there are a lot of metrics that could be chosen to define the stopping distance: type of brakes, type of tires, wet/dry surface, driver’s attention level, etc. But in this case I discovered that, based on the info provided, just using a single independent variable, with a simple linear model, I was able to explain quite well the changes in the dependent variable. 







